Motivating Exploration in Reinforcement Learning

### 2020-04-16 paper/main.tex
Updated impl details

\subsection{Intrinsic Curiosity}
Entropy encourages the agent to explore but in a purely random manner, which has been shown to produce exceptional results (as below). However, we also compare with a more targeted approach by leveraging the Intrinsic Curiosity Module (ICM) introduced in \textbf{cite}, which provides an intrinsic reward as a learning signal. ICM consists of two linear networks referred to as the forward and inverse models. The forward model $f_{Fwd}$ is trained to predict the next state ($\phi(s_{t+1})$) given the current state ($\phi(s_t)$) and the action ($a_t$) decided by the policy net. Note that we use $\phi$ to denote a feature extractor network used to encode the raw states.
\begin{equation}\label{eq:intrinsic}
r_i = \sum_t (\phi(s_{t+1}) - f_{Fwd}(\phi(s_t);a_t))^2
\end{equation}
\begin{equation}\label{eq:reward}
r = r_e + \beta r_i
\end{equation}
The new intrinsic reward $r_i$ is defined as the mean-squared-error of the predicted next state and the actual next state, which is then combined with the extrinsic reward to obtain the total reward used in the policy update \ref{eq:policy}. Thus, our agent learns to maximize the likelihood of following trajectories that maximize the intrinsic reward, which means that the agent is encouraged to enter new, unseen states that $f_{Fwd}$ is unable to predict accurately. The forward model itself is trained using the loss below.
\begin{equation}\label{eq:forward}
\mathcal{L}_{Fwd} = \frac{1}{n} \sum_t (\phi(s_{t+1}) - f_{Fwd}(\phi(s_t);a_t))^2
\end{equation}
Further, the purpose of the inverse model $f_{Inv}$ is to train the feature extractor $\phi$ to encode observed features from the raw state that are most impacted by the agent's actions. Specifically, the inverse model takes the current and next states and infers the corresponding action taken: $f_{Inv}(\phi(s_t), \phi(s_{t+1})) = \hat{p}$, where $\hat{p}$ is a vector of probabilities for each possible action. This is trained with the following,
\begin{equation}\label{eq:inverse}
\mathcal{L}_{Inv} = -\log \Big(\frac{e^{\hat{p}[a_t]}}{\sum_j e^{\hat{p}[j]}}\Big) = -\hat{p}[a_t] + \log\Big(\sum_j e^{\hat{p}[j]}\Big)
\end{equation}
The described curiosity model is incorporated into the base A2C model and can be trained jointly:
\begin{equation}
\mathcal{L} = \mathcal{L}_P + \lambda_{V}\mathcal{L}_V - \lambda_{E}\mathcal{L}_E + \mathcal{L}_{Fwd} + \mathcal{L}_{Inv}
\end{equation}

@@

\subsection{Implementation Details}

\subsection{Quantitative Results}

@@

\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Figures}

\begin{figure}[h]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\section{Tables}

\begin{table}[h]
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

\subsection{Intrinsic Curiosity}
Entropy encourages the agent to explore but in a purely random manner, which has been shown to produce exceptional results (as below). However, we also compare with a more targeted approach by leveraging the Intrinsic Curiosity Module (ICM) introduced in \textbf{cite}, which provides an intrinsic reward as a learning signal. ICM consists of two linear networks referred to as the forward and inverse models. The forward model $f_{Fwd}$ is trained to predict the next state ($\phi(s_{t+1})$) given the current state ($\phi(s_t)$) and the action ($a_t$) decided by the policy net. Note that we use $\phi$ to denote a feature extractor network used to encode the raw states.
\begin{equation}\label{eq:intrinsic}
r_i = \sum_t (\phi(s_{t+1}) - f_{Fwd}(\phi(s_t);a_t))^2
\end{equation}
\begin{equation}\label{eq:reward}
r = r_e + \beta r_i
\end{equation}
The new intrinsic reward $r_i$ is defined as the mean-squared-error of the predicted next state and the actual next state, which is then combined with the extrinsic reward to obtain the total reward used in the policy update \ref{eq:policy}. Thus, our agent learns to maximize the likelihood of following trajectories that maximize the intrinsic reward, which means that the agent is encouraged to enter new, unseen states that $f_{Fwd}$ is unable to predict accurately. The forward model itself is trained using the loss below.
\begin{equation}\label{eq:forward}
\mathcal{L}_{Fwd} = \frac{1}{n} \sum_t (\phi(s_{t+1}) - f_{Fwd}(\phi(s_t);a_t))^2
\end{equation}
Further, the purpose of the inverse model $f_{Inv}$ is to train the feature extractor $\phi$ to encode observed features from the raw state that are most impacted by the agent's actions. Specifically, the inverse model takes the current and next states and infers the corresponding action taken: $f_{Inv}(\phi(s_t), \phi(s_{t+1})) = \hat{p}$, where $\hat{p}$ is a vector of probabilities for each possible action. This is trained with the following,
\begin{equation}\label{eq:inverse}
\mathcal{L}_{Inv} = -\log \Big(\frac{e^{\hat{p}[a_t]}}{\sum_j e^{\hat{p}[j]}}\Big) = -\hat{p}[a_t] + \log\Big(\sum_j e^{\hat{p}[j]}\Big)
\end{equation}
The described curiosity model is incorporated into the base A2C model and can be trained jointly:
\begin{equation}
\mathcal{L} = \mathcal{L}_P + \lambda_{V}\mathcal{L}_V - \lambda_{E}\mathcal{L}_E + \mathcal{L}_{Fwd} + \mathcal{L}_{Inv}
\end{equation}
where we use $\lambda_V=0.5$ and $\lambda_E=0.02$.

@@

\subsection{Implementation Details}
Our work is implemented in Pytorch and we iterate our agents in the Atari environments provided through OpenAI Gym \textbf{cite}, namely \texttt{PongNoFrameskip-v0}, \texttt{SeaquestNoFrameskip-v0}, and \texttt{BreakoutNoFrameskip-v0} (using consistent random seeds for each of the environments). Specifically, we train using 4 parallel environments at a time and with each state consisting of the current frame concatenated with 3 previous frames. We use the Adam optimizer \textbf{cite} with a learning rate of 0.0001, a rollout size of 5, and a varying number of update steps (due to the limited scope of the project). All of our models are trained synchronously on Nvidia GTX 1080Ti (11GB VRAM) and Tesla V100 (16GB VRAM) gpus with batch size equivalent to the number of parallel environments.

The main actor-critic network consists of a feature extractor, followed by the actor and critic headers, which are simply single linear layers that map the extracted features to the desired next-action logits and predicted discounted rewards respectively. The feature extractor consists of four stride-2 convolutional layers followed by a LSTM layer \textbf{cite}, which is needed to leverage past information. We note that the LSTM memory is cleared at the end of every episode. The encoded features are flattened to $B\times 288$ tensors before being forwarded to the actor-critic. The initial state is resized to a $48\times 48$ image before being passed to the feature extractor.

The curiosity module being used consists of a forward and inverse model as well as another feature extractor net (this one being purely convolutional). Both forward and inverse models are simple multi-layer perceptrons comprised of two stacked linear layers with 288 hidden neurons.

\subsection{Quantitative Results}
\begin{figure}[ht]\label{fig:pong-reward}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

@@
